Coevolution leaves a stronger imprint on interactions than on community structure

### 2017-05-06 manuscript/manuscript.Rmd
replies list of todos
update the MS + replies

Historically, the evidence for shared evolutionary history in taxonomically
diverse communities relied on the quantification of the degree of matching
between the phylogenies of two sets of interacting organisms, accounting for the
distributions of interactions across the phylogeny [@lege02sth]. This notion
builds on the century-old idea that extant species interact in a way similar to
the way their ancestors did [@fahr13eua; @nuis13cam; @guim11ecm]. Note that
testing these assumptions is related to, but markedly more restrictive than,
testing for phylogenetic conservatism of species' interactions [@reze07ncp;
@eklo12reh]. This is because of additional, higher-order constraints related to
the shape of both trees at *all* depths [@cave09mce; @mouq12eap], because
ancestral evolutionary innovations have a high phylogenetic inertia, and they
carry forward to extant taxa [@vale10cpr; @desd03qps; @dini08mgc]. In a way, the
true measure of phylogenetic signal of interactions should depend not only on
how they are conserved within the tree of the species establishing them (*e.g.*
parasites or pollinators), but also how these interactions at matched to the
tree of the species receiving them (*e.g.* hosts or plants). This is true
whether or not the species complex are coevolving: in fact, neutral interactions
can yield perfectly matching co-phylogenies [@pois152wc]. Consequently, many of
the systems that have been described as exhibiting significant phylogenetic
structure of interactions ultimately deviate from this last constraint, and this
can occur for a variety of factors that stem from how other species evolved and
established, lost, or maintained interactions throughout their joint
evolutionary history. Nonetheless, detecting matching phylogenies for
interacting clades indicates that their shared evolutionary history is long
standing and is therefore suggestive that their extant ecological structure is
an outcome of ancestral constraints and/or co-adaptation [@nuis15pri].

It is important to note further that discovering matching phylogenies does not
mean that coevolutionary dynamics---*sensu* @thom99rmc---took place at any time.
In fact, coevolution is not expected to necessarily result in matching
phylogenies nor are matching phylogenies only produced through coevolution
[@pois152wc]. It follows that community-level measures of phylogenetic signal,
while they *do* quantify how closely interactions are a product of phylogeny, do
not allow us to draw conclusions on coevolution. Nevertheless, interaction-level
measures are useful, in that, when expressed as the contribution of interactions
to the overall signal, they allow us to *compare* the importance of interactions
across replicated communities. Communities from the same regional pool vary
because (i) the local species pool is at best a subset of the regional species
pool and (ii) the local interactions are at best a subset of the interactions in
the regional community [@pois15swe]. This implies that (i) the phylogenetic
signal in the regional pool will be different from the signal in the local
communities, and (ii) the phylogenetic signal across local communities will
differ. Species sampling and variability of interactions, however, does no
predict (i) how the phylogenetic signal of pairwise interactions is kept or lost
at the scale of the whole community nor (ii) whether or not this variability is
related to changes in the amount of phylogenetic signal that can be detected
locally.

@@

![Distribution of co-phylogenetic matching at the network and interaction levels. **a**, Networks that have lower co-phylogenetic matching at the local or regional level are composed of interactions that on average contribute little to co-phylogenetic matching at the continental scale. co-phylogenetic matching is presented relatively to the continental level co-phylogenetic matching. Dashed lines are a cubic smoothing spline, and the two levels of the same networks are linked by solid grey lines. **b**, Overall, interactions observed at the local, regional, and continental scale have roughly equivalent contributions to co-phylogenetic matching. Probability density was smoothed using a Gaussian kernel density estimator. Raw probability densities are shown as semi-transparent bars.](figure/contributions-1.pdf){#fig:contributions width=100%}

@@

Local networks show little to no signal of co-phylogenetic matching, and the
strength of co-phylogenetic matching that can be ascribed to the interactions
between two species is a surprisingly poor predictor of how frequently they
interact. In contrast to the frequent assumption that phylogenetic structure is
a key driver of community structure [@cave09mce], these data reveal that this
impact is actually minimal at ecologically relevant spatial scales. And yet,
despite all the above, individual interactions are somehow able to maintain
their co-phylogenetic matching even when the community they are woven into does
not. Thinking more broadly, these discrepancies provide a clear roadmap for
bridging the gap between our appreciation of the role of shared evolutionary
history and its empirically measurable outcomes: network structure is the most
parsimonious *mechanism* by which coevolution proceeds, not the imprint
potential coevolution leaves on ecological communities.

>>>>

Historically, the evidence for shared evolutionary history in taxonomically
diverse communities relied on the quantification of the degree of matching
between the phylogenies of two sets of interacting organisms, accounting for the
distributions of interactions across the phylogeny [@lege02sth]. This notion
builds on the century-old idea that extant species interact in a way similar to
the way their ancestors did [@fahr13eua; @nuis13cam; @guim11ecm]. Note that
testing these assumptions is related to, but markedly more restrictive than,
testing for phylogenetic conservatism of species' interactions [@reze07ncp;
@eklo12reh]. This is because of additional, higher-order constraints related to
the shape of both trees at *all* depths [@cave09mce; @mouq12eap], because
ancestral evolutionary innovations have a high phylogenetic inertia, and they
carry forward to extant taxa [@vale10cpr; @desd03qps; @dini08mgc]. In a way, the
true measure of phylogenetic signal of interactions should depend not only on
how they are conserved within the tree of the species establishing them (*e.g.*
parasites or pollinators), but also how these interactions at matched to the
tree of the species receiving them (*e.g.* hosts or plants). This is true
whether or not the species complex are coevolving: in fact, neutral interactions
can yield perfectly matching co-phylogenies [@pois15wce]. Consequently, many of
the systems that have been described as exhibiting significant phylogenetic
structure of interactions ultimately deviate from this last constraint, and this
can occur for a variety of factors that stem from how other species evolved and
established, lost, or maintained interactions throughout their joint
evolutionary history (in addition to the fact that species may have different
impacts on one another, *i.e.* the effect of any interaction on the
community-level cophylogenetic signal is not expected to be equal). Nonetheless,
detecting matching phylogenies for interacting clades indicates that their
shared evolutionary history is long standing and is therefore suggestive that
their extant ecological structure is an outcome of ancestral constraints and/or
co-adaptation [@nuis15pri].

It is important to note further that discovering matching phylogenies does not
mean that coevolutionary dynamics---*sensu* @thom99rmc---took place at any time.
In fact, coevolution is not expected to necessarily result in matching
phylogenies nor are matching phylogenies only produced through coevolution
[@pois15wce]. It follows that community-level measures of phylogenetic signal,
while they *do* quantify how closely interactions are a product of phylogeny, do
not allow us to draw conclusions on coevolution. Nevertheless, interaction-level
measures are useful, in that, when expressed as the contribution of interactions
to the overall signal, they allow us to *compare* the importance of interactions
across replicated communities. Communities from the same regional pool vary
because (i) the local species pool is at best a subset of the regional species
pool and (ii) the local interactions are at best a subset of the interactions in
the regional community [@pois15swe]. This implies that (i) the phylogenetic
signal in the regional pool will be different from the signal in the local
communities, and (ii) the phylogenetic signal across local communities will
differ. Species sampling and variability of interactions, however, does no
predict (i) how the phylogenetic signal of pairwise interactions is kept or lost
at the scale of the whole community nor (ii) whether or not this variability is
related to changes in the amount of phylogenetic signal that can be detected
locally.

@@

![Distribution of co-phylogenetic matching at the network and interaction levels. **a**, Networks that have lower co-phylogenetic matching at the local or regional level are composed of interactions that on average contribute little to co-phylogenetic matching at the continental scale. Co-phylogenetic matching is presented relatively to the continental level co-phylogenetic matching. Dashed lines are a cubic smoothing spline, and the two levels of the same networks are linked by solid grey lines. **b**, Overall, interactions observed at the local, regional, and continental scale have roughly equivalent contributions to co-phylogenetic matching. Probability density was smoothed using a Gaussian kernel density estimator. Raw probability densities are shown as semi-transparent bars. Interaction-level contributions are unitless, and lower values represent stronger signal.](figure/contributions-1.pdf){#fig:contributions width=100%}

@@

The lack of a relationship between interaction probability (which represents
their ecological stability) and their contribution to the cophylogenetic process
(which can be a proxy of their evolutionary stability) draws into the question
of the long-term stability of interactions. Specifically, the existence of
interactions with strong cophylogenetic contribution, but low probability, is
puzzling. It is tempting to hypothesize that these interactions were once strong
enough to drive (co-)speciation, but are now decaying due either to the fact
that the host/parasite pair has been disrupted, or under the action of
ecological processes. @janz11err show that parasites can frequently acquire
hosts that are not closely related to their initial host. This phenomenon,
dubbed the "parasite paradox", can be responsible for a large number of
diversification events, as parasites will be introduced to a novel environment
when exposed to a novel host. The acquisition of novel hosts proceeds primarily
from ecological contacts between distantly related species. This emphasizes the
need to account for the spatial structure of species interactions in these
studies -- because all species do not co-occur at all places, the potential for
ecological contact with novel hosts varies across space. Ideally, reconstruction
of the ancestral range of the species can shed some light on whether acquisiton
of distantly related species happened, and where.

Local networks show little to no signal of co-phylogenetic matching, and the
strength of co-phylogenetic matching that can be ascribed to the interactions
between two species is a surprisingly poor predictor of how frequently they
interact. In contrast to the frequent assumption that phylogenetic structure is
a key driver of community structure [@cave09mce], these data reveal that this
impact is actually minimal at ecologically relevant spatial scales, even in
system where phylogenetic structure would be expected to have a profoundly
strong role. And yet, despite all the above, individual interactions are somehow
able to maintain their co-phylogenetic matching even when the community they are
woven into does not. Thinking more broadly, these discrepancies provide a clear
roadmap for bridging the gap between our appreciation of the role of shared
evolutionary history and its empirically measurable outcomes: network structure
is the most parsimonious *mechanism* by which coevolution proceeds, not the
imprint potential coevolution leaves on ecological communities.

Finally, we would like to acknowledge some limitations in the data.
Host-parasite evolutions usually follows a number of canonical coevolutionary
scenarios [@page03ttp]: true cospeciation, wherein the two speciation events are
simultaneous; pseudo- or delayed-cospeciations, wherein one of the speciations
happens after the other; host-shift, wherein the parasite acquires a new host
(and can subsequently lose the previous one, undergo speciation, or maintain its
ability to use it); finally, host range expansion, in which the parasite
acquires a novel host. When the likelihood of both host and parasite extinctions
events are added to these model, the variety of outcomes is considerable.
@dev07wch shows that host-switch like events can result in matching trees, and
so one could argue that the continental-level signal of tree matching can emerge
from numerous host-shift events [although the results of @kras12psm do not
support this scenario]. Because separating true cospeciation events from all
other situations is difficult without time-calibrated trees, @dev13cvh suggest
that the number of cospeciation event is inflated by cophenetic methods like
*PACo* or *ParaFit* [although @lope01mpf reported an increased number of
cospeciations when accounting for time of speciation events]. This re-inforces
the need for high-quality phylogeny data calibrated for communities in which the
interactions are known. Not only will this help make more robust inference about
the evolution of networks [@eklo16pcf], it will also help use informations on
species interactions embedded within networks to help understand coevoluationary
dynamics.
